---
title: "<small>Module R’Stat1 : Rappels (2/4)</small>"
author: <small><francois.rebaudo@ird.fr></small>
institute: IRD
date: Novembre 2019 ; IRD-Montpellier-France <p style="text-align:center"><small>CC BY-NC-ND 3.0</small></p>
subtitle: ""
---

# Rappels : les conteneurs de données - `vector` {data-background=#273142}

## `vector`

```{r}
miVec01 <- c(1, 2, 3, 4)
```

## `vector`

```{r}
miVec01[2]
miVec01[2:4]
miVec01[c(1, 1, 3, 2)]
```

## `vector`

```{r}
miVec01[miVec01 > 2]
miVec01[miVec01 > 2 & miVec01 < 4]
miVec01[miVec01 == 1]
```

## `vector`

```{r}
miVec01 <- c(10, 20, 30, 40, 50)
miVec02 <- c("A", "B", "C", "D", "E")
miVec02[miVec01 > 20]
```

## `vector`

```{r}
miVec02[-1]
miVec02[-c(1, 2, 5)]
```

## `vector`

```{r}
miVec02 <- c("A", "B", "C", "D", "E")
miVec03 <- c(TRUE, TRUE, FALSE, TRUE, FALSE)
miVec02[miVec03]
```

## `vector`

```{r}
miVec04 <- c(aaa = 10, bbb = 20, ccc = 30, ddd = 40, eee = 50)
miVec04["bbb"]
miVec04[c("bbb", "ccc", "bbb")]
names(miVec04)
```

## `vector`

```{r}
miVec20 <- c(10, 20, 30, 40, 50, 60, 70, 80, 90)
miVec21 <- c("a", "b", "c", "d", "e", "f", "g", "h", "i")
miVec21[3] <- "zzz"
print(miVec21)
miVec21[(miVec20 >= 50) & ((miVec20 < 80))] <- "qwerty"
print(miVec21)
```

## `vector`

```{r}
names(miVec04)[2] <- "bb_bb"
```

# Rappels : les conteneurs de données - `list` {data-background=#273142}

## `list`

```{r}
miList01 <- list()
print(miList01)
miList02 <- list(5, "qwerty", c(4, 5, 6), c("a", "b", "c"))
print(miList02)
```

## `list`

```{r}
miList03 <- list(5, "qwerty", list(c(4, 5, 6), c("a", "b", "c")))
print(miList03)
```

## `list`

```{r}
print(miList02)
miList02[[1]]
```

## `list`

```{r}
print(miList02)
miList02[[4]][2]
```

## `list`

```{r}
miList04 <- list(miList02, miList03)
print(miList04)
```

## `list`

```{r}
miList04[[2]][[3]][[1]][1]
```

## `list`

```{r}
bddInsect <- list(
  Snonagrioides = list(
    site01 = c(12, 5, 8, 7), 
    site02 = c(5, 23, 4, 41), 
    site03 = c(12, 0, 0, 0)
  ), 
  Onubilalis = list(
    site01 = c(12, 1, 2, 3), 
    site02 = c(0, 0, 0, 1), 
    site03 = c(1, 1, 2, 3)
  )
)
```

## `list`

```{r}
print(bddInsect)
```

## `list`

```{r}
bddInsect[[1]][[2]]
bddInsect$Snonagrioides$site02
bddInsect[['Snonagrioides']][['site02']]
```

## `list`

```{r}
names(bddInsect)
names(bddInsect[[1]])
```

## `list`

```{r}
bddInsect[[1]][[2]]
is.list(bddInsect[[1]][[2]])
bddInsect[[1]][2]
is.list(bddInsect[[1]][2])
```

## `list`

```{r}
bddInsect[[1]][c(1, 2)]
bddInsect[['Snonagrioides']][c("site01", "site02")]
```

## `list`

```{r}
bddInsect[['Snonagrioides']][['site02']] <- c(2, 4, 6, 8)
print(bddInsect)
```

## `list`

```{r}
miList05 <- list("a", c("b", "c"), "d")
print(miList05)
unlist(miList05)
```

# Rappels : les conteneurs de données - `data.frame` {data-background=#273142}

## `data.frame`

```{r}
miDf01 <- data.frame(
  numbers = c(1, 2, 3, 4), 
  logicals = c(TRUE, TRUE, FALSE, TRUE), 
  characters = c("a", "b", "c", "d")
)
print(miDf01)
```

## `data.frame`

```{r}
numbers <- c(1, 2, 3, 4)
logicals <- c(TRUE, TRUE, FALSE, TRUE)
characters <- c("a", "b", "c", "d")
miDf01 <- data.frame(numbers, logicals, characters)
print(miDf01)
```

## `data.frame`

```{r}
str(miDf01)
```

## `data.frame`

```{r}
miDf01 <- data.frame(
  numbers = c(1, 2, 3, 4), 
  logicals = c(TRUE, TRUE, FALSE, TRUE), 
  characters = c("a", "b", "c", "d"),
  stringsAsFactors = FALSE
)
```

## `data.frame`

```{r}
str(miDf01)
```

## `data.frame`

```{r}
miDf01 <- data.frame(
  numbers = c(1, 2, 3, 4), 
  logicals = c(TRUE, TRUE, FALSE, TRUE), 
  characters = c("a", "b", "c", "d")
)
miDf01$characters <- as.character(miDf01$characters)
```

## `data.frame` {data-background=#842a1d}

<h1> TOUJOURS vérifier ces données avec `str()` <h1>

## `data.frame`

```{r}
miDf01$numbers # vector
miDf01[[1]] # vector
miDf01[["numbers"]] # vector
```

## `data.frame`

```{r}
miDf01[1] # list
miDf01["numbers"] # list
```

## `data.frame`

```{r}
myRow <- 2
myCol <- 1
miDf01[myRow, myCol]
```

## `data.frame`

```{r}
miDf01[myRow, ]
miDf01[, myCol]
miDf01[, "numbers"]
```

## `data.frame`

```{r}
miDf01[, c(1, 2)]
miDf01[c(2, 1), ]
```

## `data.frame`

```{r}
miDf01[miDf01$numbers > 2, ]
miDf01[(miDf01$logicals == TRUE) & (miDf01$numbers < 2), ]
```

## `data.frame`

```{r}
miDf01$newVar <- c(4, 5, 6, 7)
print(miDf01)
```

## `data.frame`

```{r}
miDf01$numbers <- miDf01$numbers + 100
print(miDf01)
```

## `data.frame`

```{r}
colnames(miDf01)
rownames(miDf01)
```

## `data.frame`

```{r}
miDf01 <- rbind(miDf01, 
  nouvelleLigne = data.frame(
    numbers = 5, 
    logicals = TRUE, 
    characters = "e", 
    newVar = 8))
```

## `data.frame`

```{r}
rownames(miDf01)
rownames(miDf01) <- c("R1", "R2", "R3", "R4", "R5")
```

## `data.frame`

```{r}
print(miDf01)
```

# Rappels : les conteneurs de données - `matrix` {data-background=#273142}

## `matrix`

```{r}
vecForMatrix <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)
miMat <- matrix(vecForMatrix, nrow = 3, ncol = 4)
print(miMat)
```

## `matrix`

```{r}
vecForMatrix <- 1:12
miMat <- matrix(vecForMatrix, nrow = 3)
print(miMat)
```

## `matrix`

```{r}
miMat <- matrix(
  vecForMatrix, 
  nrow = 3, 
  byrow = TRUE)
print(miMat)
```

## `matrix`

```{r}
miMat <- matrix(
  vecForMatrix, 
  nrow = 3, 
  byrow = TRUE, 
  dimnames = list(
    c("r1", "r2", "r3"), 
    c("c1", "c2", "c3", "c4"))
)
print(miMat)
```

## `matrix`

```{r}
miMat[c(1, 2), c(2, 3)]
miMat[10]
```

# Rappels : les conteneurs de données - `array` {data-background=#273142}

## `array` 

```{r}
miMat01 <- matrix(1:12, ncol = 3)
dim(miMat01)
```

## `array` 

```{r}
miVecArr <- 1:24
miArray <- array(miVecArr, dim = c(3, 4, 2))
print(miArray)
```

## `array` 

```{r}
miArray02 <- array(1:24, 
  dim = c(3, 4, 2), 
  dimnames = list(
    c("r1", "r2", "r3"), 
    c("c1", "c2", "c3", "c4"), 
    c("matrix1", "matrix2")
  )
)
print(miArray02)
```

## `array` 

```{r}
i <- 2
j <- 1
k <- 1
miArray02[i, j, k]
miArray02[, j, k]
```

## `array` 

```{r}
miArray02[i, , k]
miArray02[i, j, ]
```


# <a href = "R013_intro.html"> SUIVANT </a> {data-background=#273142}
