---
title: "<small>Module R’Stat1 : Rappels (3/4)</small>"
author: <small><francois.rebaudo@ird.fr></small>
institute: IRD
date: Novembre 2019 ; IRD-Montpellier-France <p style="text-align:center"><small>CC BY-NC-ND 3.0</small></p>
subtitle: ""
---
  
# Rappels : les graphiques simples - `plot` {data-background=#273142}

## `plot`

```{r, fig.width = 10, fig.height = 5}
myX <- rnorm(100, mean = 0, sd = 1)
myY <- rnorm(100, mean = 10, sd = 1)
plot(x = myX, y = myY)
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
plot(x = myX, y = myY, 
  xlab = "XXXXXXXXXX", ylab = "YYYYYYYYYY")
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
plot(x = myX, y = myY, 
  xlim = c(-10, 10), ylim = c(5, 15))
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
plot(x = rep(1:5, 5), y = rep(1:5, each = 5),
  pch = 1:25, xlab = "", ylab = "", cex = 2)
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
plot(x = rep(1:5, 5), y = rep(1:5, each = 5),
  pch = 1:25, xlab = "", ylab = "", cex = 2)
text(x = rep(seq(1:5), 5), y = rep(seq(1:5), each = 5),
  labels = 1:25, pos = 4)
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
plot(x = myX, y = myY, 
  pch = c("a", "@", "#", "1", "=", "-", "_", "o", "O", "0", 
    letters[1:15]), cex = 2)
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
plot(x = myX, y = myY, 
  cex = seq(from = 0.5, to = 5, length.out = 100))
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
plot(x = myX, y = myY, 
  cex = seq(from = 0.5, to = 5, length.out = 100),
  pch = 16,
  col = sample(colors(), 100))
```

## `plot`

```{r}
myX <- rnorm(100)
myY <- rnorm(100)
dfGraph <- data.frame(myX, myY)
```

## `plot`

```{r}
head(dfGraph, n = 20)
```

## `plot`

```{r}
dfGraph <- dfGraph[order(dfGraph$myX),]
dfGraph$myCol <- colorRampPalette(c("blue", "yellow", "red"))(100)
```

## `plot`

```{r}
head(dfGraph, n = 20)
```

## `plot`

```{r}
dfGraph <- dfGraph[order(dfGraph$myY),]
dfGraph$myCex <- seq(from = 0.5, to = 5, length.out = 100)
```

## `plot`

```{r}
head(dfGraph, n = 20)
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
plot(x = dfGraph$myX, y = dfGraph$myY, 
  cex = dfGraph$myCex, pch = 16, col = dfGraph$myCol, 
  xlab = "", ylab = "")
```

## `plot`

```{r}
dfGraph$grupo <- factor(
  sample(letters[1:5], size = 100, replace = TRUE))
```

## `plot`

```{r}
head(dfGraph, n = 20)
```

## `plot`

```{r}
str(dfGraph)
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
plot(x = dfGraph$myX, y = dfGraph$myY, 
  pch = 16, xlab = "", ylab = "",
  col = as.numeric(dfGraph$grupo))
```

## `plot`

```{r, fig.width = 10, fig.height = 4.5}
plot(x = dfGraph$myX, y = dfGraph$myY, pch = 16, xlab = "", 
  ylab = "", cex = 2, col = as.numeric(dfGraph$grupo))
legend("topleft", fill = seq_along(levels(dfGraph$grupo)), 
  legend = levels(dfGraph$grupo))
```

## `plot`

```{r, fig.width = 10, fig.height = 4}
plot(x = dfGraph$myX, y = dfGraph$myY, 
  pch = as.numeric(dfGraph$grupo), xlab = "", ylab = "", 
  cex = 2, col = as.numeric(dfGraph$grupo))
legend("topleft", pch = seq_along(levels(dfGraph$grupo)), 
  legend = levels(dfGraph$grupo), 
  col = seq_along(levels(dfGraph$grupo)))
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
myX <- 1:20
myY <- rnorm(20, mean = 10, sd = 1)
plot(x = myX, y = myY, type = 'b')
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
myX <- 1:20
myY <- rnorm(20, mean = 10, sd = 1)
plot(x = myX, y = myY, type = 'p') 
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
myX <- 1:20
myY <- rnorm(20, mean = 10, sd = 1)
plot(x = myX, y = myY, type = 'l') 
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
myX <- 1:20
myY <- rnorm(20, mean = 10, sd = 1)
plot(x = myX, y = myY, type = 'c') 
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
myX <- 1:20
myY <- rnorm(20, mean = 10, sd = 1)
plot(x = myX, y = myY, type = 'o') 
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
myX <- 1:20
myY <- rnorm(20, mean = 10, sd = 1)
plot(x = myX, y = myY, type = 'h') 
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
myX <- 1:20
myY <- rnorm(20, mean = 10, sd = 1)
plot(x = myX, y = myY, type = 's') 
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
myX <- 1:20
myY <- rnorm(20, mean = 10, sd = 1)
plot(x = myX, y = myY, type = 'n') 
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
plot(x = myX, y = myY, type = 'n') 
grid(lwd = 3, lty = 1)
points(x = myX, y = myY, type = "l", lwd = 2)
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
plot(x = myX, y = myY, type = 'l', lwd = 2) 
grid(lwd = 3, lty = 1)
```

## `plot`

```{r, fig.width = 10, fig.height = 5}
plot(x = myX, y = myY, type = 'l', lwd = 2, 
  panel.first = grid(lwd = 3, lty = 1)) 
```

## `plot`

```{r, fig.width = 10, fig.height = 4}
op <- par(no.readonly = TRUE)
par(mfrow = c(1, 2))
plot(x = myX, y = myY, type = 'l', lwd = 2) 
grid(lwd = 3, lty = 1)
plot(x = myX, y = myY, type = 'l', lwd = 2, 
  panel.first = grid(lwd = 3, lty = 1))
```

## `plot`

```{r}
par(op)
```

## `plot`

```{r, fig.width = 10, fig.height = 4}
par(mfrow = c(1, 2), mar = c(2, 2, 1, 1))
plot(x = myX, y = myY, type = 'l', lwd = 2) 
grid(lwd = 3, lty = 1)
plot(x = myX, y = myY, type = 'l', lwd = 2, 
  panel.first = grid(lwd = 3, lty = 1))
```

## `plot`

```{r}
par(op)
```

## `plot`

```{r, fig.width = 10, fig.height = 4}
par(mfrow = c(2, 2), mar = c(2, 2, 1, 1))
graph4 <- lapply(1:4, function(i){
  plot(x = rnorm(100), 
    y = rnorm(100), 
    col = i, pch = 16)
})
```

## `plot`

```{r}
par(op)
```

## `plot`

```{r, fig.width = 10, fig.height = 4}
plot(x = myX, y = myY, type = 'l', lwd = 2)
abline(h = mean(myY), lwd = 1, lty = 2)
abline(v = myX[myY == min(myY)], col = 4, lty = 2)
abline(v = myX[myY == max(myY)], col = 2, lty = 2)
```

# Rappels : les graphiques simples - `hist` {data-background=#273142}

## `hist`

```{r}
myX <- list(
  rnorm(1000),
  rgamma(1000, shape = 1),
  sample(1:100, size = 1000, replace = TRUE),
  rbeta(1000, shape1 = 1, shape2 = 2)
)
myTitle <- c("Normal", "Gamma", "Uniform", "Beta")
```

## `hist`

```{r, fig.width = 10, fig.height = 4}
par(mfrow = c(2, 2), mar = c(2, 2, 1, 1))
tr <- lapply(1:4, function(i){
  hist(myX[[i]], col = heat.colors(15), 
    main = myTitle[i]
  )
})
```

## `hist`

```{r}
par(op)
```

## `hist`

```{r, fig.width = 10, fig.height = 4}
myX <- rnorm(1000)
par(mfrow = c(1, 2), mar = c(1, 1, 1, 1))
hist(myX, axes = FALSE, col = rainbow(100))
hist(myX, axes = FALSE, nclass = 100, col = rainbow(100))
```

## `hist`

```{r}
par(op)
```

# Rappels : les graphiques simples - `barplot` {data-background=#273142}

## `barplot`

```{r echo = FALSE, message = FALSE}
pkgCheck <- function(x){ 
    if (!require(x, character.only = TRUE)){
        install.packages(x, dependencies = TRUE)
        if(!require(x, character.only = TRUE)) {
            stop()
        }
    }
}
pkgCheck("palettesForR")
```

## `barplot`

```{r eval = FALSE}
pkgCheck <- function(x){ 
    if (!require(x, character.only = TRUE)){
        install.packages(x, dependencies = TRUE)
        if(!require(x, character.only = TRUE)) {
            stop()
        }
    }
}
pkgCheck("palettesForR")
```

## `barplot`

```{r, fig.width = 10, fig.height = 4.5}
# pkgCheck("palettesForR")
myX <- rnorm(3, mean = 10)
barplot(myX, names.arg = c("A", "B", "C"), col = Cool_gpl)
```

## `barplot`

```{r, fig.width = 10, fig.height = 5}
par(mar = c(1, 1, 1, 1))
myX <- matrix(c(4, 5, 8, 4, 6, 2, 3, 4, 5), nrow = 3)
barplot(myX, axes = FALSE, col = Pastels_gpl[1:3])
```

## `barplot`

```{r, fig.width = 10, fig.height = 5}
par(mar = c(1, 1, 1, 1))
barplot(myX, axes = FALSE, col = Pastels_gpl[1:3], 
  beside = TRUE, legend.text = c("A", "B", "C"))
```

## `plot` + `barplot`

```{r, echo = FALSE}
n <- 50
myX <- rnorm(n)
myY <- rnorm(n)
myCol <- c(2, sample(c(4, 2), size = (n - 1), replace = TRUE))
myColors <- colorRampPalette(c("blue", "red"))(100)
myYCut <- cut(myY, breaks = -4:4)
myXCut <- cut(myX, breaks = -4:4)
myYCutCol <- table(myCol, myYCut)
myXCutCol <- table(myCol, myXCut)
xCol <- round(
  myXCutCol[1,] / (myXCutCol[1,] + myXCutCol[2,]) * 100
)
xCol[is.na(xCol) | xCol == 0] <- 1
yCol <- round(
  myYCutCol[1,] / (myYCutCol[1,] + myYCutCol[2,]) * 100
)
yCol[is.na(yCol) | yCol == 0] <- 1
par(mar = c(2, 3, 1, 1))
layout(matrix(c(1, 1, 0, 
                2, 2, 3, 
                2, 2, 3), ncol = 3, byrow = TRUE))
barplot(table(myXCut), las = 1, col = myColors[xCol])
plot(x = myX, y = myY, col = myCol, pch = 16, 
  xlim = c(-4, 4), ylim = c(-4, 4), cex = 1.5, 
  panel.first = grid())
barplot(table(myYCut), las = 1, horiz = TRUE, 
  col = myColors[yCol])
```

## `plot` + `barplot`

```{r}
n <- 50
myX <- rnorm(n)
myY <- rnorm(n)
myCol <- c(2, sample(c(4, 2), size = (n - 1), replace = TRUE))
myColors <- colorRampPalette(c("blue", "red"))(100)
head(data.frame(myX, myY, myCol))
```

## `plot` + `barplot`

```{r}
myYCut <- cut(myY, breaks = -4:4)
myXCut <- cut(myX, breaks = -4:4)
print(myXCut)
```

## `plot` + `barplot`

```{r}
myYCutCol <- table(myCol, myYCut)
myXCutCol <- table(myCol, myXCut)
print(myXCutCol)
```

## `plot` + `barplot`

```{r}
xCol <- round(
  myXCutCol[1,] / (myXCutCol[1,] + myXCutCol[2,]) * 100
)
xCol[is.na(xCol) | xCol == 0] <- 1
yCol <- round(
  myYCutCol[1,] / (myYCutCol[1,] + myYCutCol[2,]) * 100
)
yCol[is.na(yCol) | yCol == 0] <- 1
print(xCol)
```

## `plot` + `barplot`

```{r eval = FALSE}
par(mar = c(2, 3, 1, 1))
layout(matrix(c(1, 1, 0, 
                2, 2, 3, 
                2, 2, 3), ncol = 3, byrow = TRUE))
barplot(table(myXCut), las = 1, col = myColors[xCol])
plot(x = myX, y = myY, col = myCol, pch = 16, 
  xlim = c(-4, 4), ylim = c(-4, 4), cex = 1.5, 
  panel.first = grid())
barplot(table(myYCut), las = 1, horiz = TRUE, 
  col = myColors[yCol])
```

## `plot` + `barplot`

```{r echo = FALSE}
par(mar = c(2, 3, 1, 1))
layout(matrix(c(1, 1, 0, 
                2, 2, 3, 
                2, 2, 3), ncol = 3, byrow = TRUE))
barplot(table(myXCut), las = 1, col = myColors[xCol])
plot(x = myX, y = myY, col = myCol, pch = 16, 
  xlim = c(-4, 4), ylim = c(-4, 4), cex = 1.5, 
  panel.first = grid())
barplot(table(myYCut), las = 1, horiz = TRUE, 
  col = myColors[yCol])
```

## `plot` + `barplot`

```{r}
graphBarplotCol <- function(n){
  myX <- rnorm(n)
  myY <- rnorm(n)
  myCol <- c(2, sample(c(4, 2), size = (n - 1), replace = TRUE))
  myColors <- colorRampPalette(c("blue", "red"))(100)
  myYCut <- cut(myY, breaks = -4:4)
  myXCut <- cut(myX, breaks = -4:4)
  myYCutCol <- table(myCol, myYCut)
  myXCutCol <- table(myCol, myXCut)
  xCol <- round(
    myXCutCol[1,] / (myXCutCol[1,] + myXCutCol[2,]) * 100
  )
  xCol[is.na(xCol) | xCol == 0] <- 1
  yCol <- round(
    myYCutCol[1,] / (myYCutCol[1,] + myYCutCol[2,]) * 100
  )
  yCol[is.na(yCol) | yCol == 0] <- 1
  op <- par(no.readonly = TRUE)
  par(mar = c(2, 3, 1, 1))
  layout(matrix(c(1, 1, 0, 
                  2, 2, 3, 
                  2, 2, 3), ncol = 3, byrow = TRUE))
  barplot(table(myXCut), las = 1, col = myColors[xCol])
  plot(x = myX, y = myY, col = myCol, pch = 16, 
    xlim = c(-4, 4), ylim = c(-4, 4), cex = 1.5, 
    panel.first = grid())
  barplot(table(myYCut), las = 1, horiz = TRUE, 
    col = myColors[yCol])
  par(op)
}
```

## `plot` + `barplot`

```{r eval = FALSE}
graphBarplotCol(n = 10)
```

## `plot` + `barplot`

```{r echo = FALSE}
graphBarplotCol(n = 10)
```

## `plot` + `barplot`

```{r eval = FALSE}
graphBarplotCol(n = 1000)
```

## `plot` + `barplot`

```{r echo = FALSE}
graphBarplotCol(n = 1000)
```

# Rappels : les graphiques simples - `boxplot` {data-background=#273142}

## `boxplot`

```{r}
df <- data.frame(
  box1 = rnorm(1000), 
  box2 = rgamma(1000, shape = 1), 
  box3 = sample(-3:3, size = 1000, replace = TRUE),
  box4 = rbeta(1000, shape1 = 1, shape2 = 2)
)
```

## `boxplot`

```{r eval = FALSE}
boxplot(df, col = c(
  rgb(0, 94, 255, maxColorValue = 255),  
  rgb(255, 0, 174, maxColorValue = 255),  
  rgb(255, 136, 0, maxColorValue = 255),  
  rgb(119, 255, 0, maxColorValue = 255)))
```

## `boxplot`

```{r echo = FALSE}
boxplot(df, col = c(
  rgb(0, 94, 255, maxColorValue = 255),  
  rgb(255, 0, 174, maxColorValue = 255),  
  rgb(255, 136, 0, maxColorValue = 255),  
  rgb(119, 255, 0, maxColorValue = 255)))
```

## `boxplot`

```{r}
df$cat <- sample(
  c("w", "x", "y", "z"), 
  size = 1000, 
  replace = TRUE)
```

## `boxplot`

```{r eval = FALSE}
boxplot(df$box3 ~ df$cat, col = c(
  rgb(0, 94, 255, maxColorValue = 255),  
  rgb(255, 0, 174, maxColorValue = 255),  
  rgb(255, 136, 0, maxColorValue = 255),  
  rgb(119, 255, 0, maxColorValue = 255)), ylab = "Box3")
```

## `boxplot`

```{r echo = FALSE}
boxplot(df$box3 ~ df$cat, col = c(
  rgb(0, 94, 255, maxColorValue = 255),  
  rgb(255, 0, 174, maxColorValue = 255),  
  rgb(255, 136, 0, maxColorValue = 255),  
  rgb(119, 255, 0, maxColorValue = 255)), ylab = "Box3")
```

## `boxplot`

```{r}
df$cat2 <- sample(
  1:3, 
  size = 1000, 
  replace = TRUE)
```

## `boxplot`

```{r eval = FALSE}
boxplot(df$box4 ~ df$cat*df$cat2, col = c(
  rgb(0, 94, 255, maxColorValue = 255),  
  rgb(255, 0, 174, maxColorValue = 255),  
  rgb(255, 136, 0, maxColorValue = 255),  
  rgb(119, 255, 0, maxColorValue = 255)), ylab = "Box4")
```

## `boxplot`

```{r echo = FALSE}
boxplot(df$box4 ~ df$cat*df$cat2, col = c(
  rgb(0, 94, 255, maxColorValue = 255),  
  rgb(255, 0, 174, maxColorValue = 255),  
  rgb(255, 136, 0, maxColorValue = 255),  
  rgb(119, 255, 0, maxColorValue = 255)), ylab = "Box4")
```

## `boxplot`

```{r eval = FALSE}
boxplot(df$box4 ~ df$cat2*df$cat, col = c(
  rgb(0, 94, 255, maxColorValue = 255),  
  rgb(255, 0, 174, maxColorValue = 255),  
  rgb(255, 136, 0, maxColorValue = 255)), ylab = "Box4")
```

## `boxplot`

```{r echo = FALSE}
boxplot(df$box4 ~ df$cat2*df$cat, col = c(
  rgb(0, 94, 255, maxColorValue = 255),  
  rgb(255, 0, 174, maxColorValue = 255),  
  rgb(255, 136, 0, maxColorValue = 255)), ylab = "Box4")
```

# <a href = "R016_intro_TD02.html"> TD02 </a> {data-background=#273142}

# <a href = "R014_intro.html"> SUIVANT </a> {data-background=#273142}




